```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RColorBrewer)
library(moosefun)
source("DetermineK.R")

pfile = "parameters/pfile.txt" 
pardata = read.table(pfile, header = FALSE, sep = "=", as.is = TRUE)

```

## Read in and format the data

```{r}
w = which(pardata[,1] == "RESULTSDIR")
results_dir = pardata[w,2]
smartpca_dir = paste0(results_dir, "smartpca/")
## Identify smartPCA files
files = list.files(smartpca_dir)
w = grep("eigen", files)
files = files[w]
## Population IDs
pops = c("AFR","EUR","SAS","EAS")
# pops = c("AFR", "SAS","EAS")
## Read in the data
mydata = lapply(pops, function(pop){
    ## Read in the eigenvalues and estimate the variance explained
    f = paste0(smartpca_dir, pop, ".eigenvalues")
    varexp = read.table( f, header = FALSE, as.is = TRUE )[,1]
    varexp = varexp/sum(varexp) *100
    ## Read in the eigenvectors or PCs
    f = paste0(smartpca_dir, pop, ".eigenvectors")
    pcs = read.table( f, header = FALSE, as.is = TRUE)
    colnames(pcs) = c("ID", paste0("PC", 1:100), "pop")
    pcs$pop = gsub("UKBB_relatives","UKBB", pcs$pop )
    ###
    pop_order = sort(unique(pcs$pop))
    w = which(pop_order == "UKBB")
    pop_order = pop_order[-w]
    pop = factor(pcs$pop, levels = c("UKBB", pop_order))
    pcs$pop = pop
    ### reorder data frame for UKBB first
    w = grep("UKBB", pop)
    pcs = rbind(pcs[w, ], pcs[-w,])
    ### add color
    cols = RColorBrewer::brewer.pal( length(unique(pop_order)) , "Set1")
    cols = c("grey",cols)
    pcs$pcol = cols[ as.factor( pcs$pop ) ]
    ## add shapes
    pcs$ppch =  c(1, rep(20,length(pop_order)) )[pcs$pop]
    ## add shape size
    pcs$pcex =  c(1, rep(2,length(pop_order)) )[pcs$pop]
    ## 
    out = list(pcs = pcs, varexp = varexp)
    return(out)

})
names(mydata) = pops

```

## plot the scree plots for the PCs

```{r, fig.width = 6, fig.height = 3}
figures_dir = paste0(results_dir, "figures/")
f = paste0(figures_dir, "plot01_Scree.pdf")
####
pdf( f , width = 8, height = 8)
par(mfrow = c(2,2), oma = c(2,1,2,0), mar = c(5,5,2,1))
for(i in 1:length(pops)){
  plot( mydata[[i]]$varexp[1:20], pch = 21, cex = 1.5, bg = "grey", type = "b",
        main = names(mydata)[i],
        xlab = "PC", ylab = "% Variance Explained")
  abline(h = 0.1, col = "black", lty = 2, lwd =2)
}
dev.off()
```

## Plot the PCs with 1KG pops

```{r, fig.width = 10, fig.height = 8}
f = paste0(figures_dir, "plot02_continental_PCs.pdf")
pdf(f, width = 8, height = 9)

par(mfrow = c(4,3), oma = c(2,1,2,0), mar = c(5,5,2,1))
for(i in 1:length(pops)){
  pop = names(mydata)[i]
  varexp = signif( mydata[[i]]$varexp[1:20] , d = 3)
  plotdata = mydata[[i]]$pcs
  plotdata = plotdata[, c("PC1","PC2","PC3","PC4", "PC5", "pop","pcol","ppch" ,"pcex")]
  for(j in 2:4){
    plot(plotdata$PC1, plotdata[,j],
         col = plotdata$pcol,
         pch = plotdata$ppch,
         cex = plotdata$pcex,
         xlab = paste0("PC1 varexp = ", varexp[1], "%" ),
         ylab = paste0("PC",j," varexp =  ", varexp[j], "%" ),
         main = pop)
    abline(h = 0, lty = 2, col = "grey", lwd = 0.75)
    abline(v = 0, lty = 2, col = "grey", lwd = 0.75)
    if(j == 3){
      ## identify the proper color for each pop
      p = levels(plotdata$pop)
      cols = sapply(p, function(x){ 
        w = which(plotdata$pop == x)[1]; return(plotdata$pcol[w]) 
        })
      ### plot the legend
      if(pop == "AFR"){
        legend("topleft", legend = levels(plotdata$pop), 
             pch = 20, col = cols, pt.cex = 2.5, cex = 0.75, bg = NULL)
      } 
      if(pop == "EUR"){
        legend("bottomright", legend = levels(plotdata$pop), 
             pch = 20, col = cols, pt.cex = 2.5, cex = 0.75, bg = NULL)
      }
      if(pop %in% c("EAS" , "SAS") ){
          legend("topright", legend = levels(plotdata$pop), 
               pch = 20, col = cols, pt.cex = 2.5, cex = 0.75, bg = NULL)  
        }
     }
  }
}

dev.off()

```


```{r, fig.width = 8, fig.height= 2}

ReferencePop_Plots = lapply(pops, function(pop){
  ####################################
  ## Define the working plot data set
  ####################################
  plotdata = mydata[[pop]]$pcs
  l = length( levels(plotdata$pop) ) -1
  plot_colors = c( "grey", RColorBrewer::brewer.pal(l, "Set1") )
  ####################################
  ## Define variance Explained
  ####################################
  varexp = mydata[[pop]]$varexp[1:10]
  varexp = round(varexp, d = 3)
  
  ################################
  ## PLOT
  ################################
  myplots = lapply(2:4, function(i){
    pc = paste0("PC", i)
    plotdata %>% ggplot( aes_string(x = "PC1", y = pc)  ) +
    geom_point( aes(color = pop), shape = 19 ) + 
    scale_color_manual( values = plot_colors, drop = FALSE ) +
    theme_bw() +
    # scale_fill_manual( values = UNregions$colors, drop = FALSE ) +
    labs(color = "Population", 
         x = paste0("PC1  ", varexp[1], "%"),
         y = paste0("PC", i, "  ", varexp[i], "%")) +
    #guides(color=guide_legend(ncol=2)) +
    guides(color=guide_legend(ncol=1,  override.aes = list(size=4) ) )
  })
  
  p = ggpubr::ggarrange(plotlist = myplots,
                    ncol=3, nrow=1,
                    common.legend = TRUE,
                    legend="right")
  myplots$combined = p
  return(myplots)
})
names(ReferencePop_Plots) = pops


####################
## Save the ggplot
## objects to 
## file
####################
f = paste0(figures_dir, "ReferencePop_Plots.Rdata")  
save(ReferencePop_Plots, file = f)


```

```{r, fig.width = 10, fig.height = 4}
myplots = list(ReferencePop_Plots[["AFR"]]$combined,
               ReferencePop_Plots[["EUR"]]$combined,
               ReferencePop_Plots[["SAS"]]$combined,
               ReferencePop_Plots[["EAS"]]$combined)

f = paste0(figures_dir, "plot02_continental_PCs_ggplot.pdf")  
pdf(f, width = 11, height = 12)
ggpubr::ggarrange(plotlist = myplots,
                    ncol=1, nrow=4,
                    labels = c("A","B","C", "D"))
dev.off()
```

## Run Kmeans

```{r}
# number_of_pcs = c(4, 4, 5, 4)
# names(number_of_pcs) = pops
# ###
# kmeans = lapply(pops, function(pop){
#   #########################
#   ## 1. how many PCs to use ?
#   #########################
#   w = which(names(number_of_pcs) == pop)
#   pcount = number_of_pcs[w]
#   #########################
#   ## 2. define data set to run
#   #########################
#   wdata = mydata[[pop]]$pcs
#   keepukbb = which(wdata$pop == "UKBB")
#   wdata = wdata[keepukbb, ]
#   p = paste0("PC", 1:pcount)
#   wdata = wdata[, p]
#   #########################
#   ## 3. run kmeans
#   #########################
#   Tstart =  Sys.time()
#   K = DetermineK(mydata = wdata, 
#              kmax = 2,
#              num_iter = 100, 
#              plotrepress =  TRUE)
#   Tend = Sys.time()
#   Tend - Tstart
#   
#   return(K)
# })
# 
# names(kmeans) = pops

f = paste0(results_dir, "kmeans/kmeans.Rdata")
load(f)

```


## Plot Silhouette

```{r, fig.width = 5, fig.height = 4}
f = paste0(results_dir, "figures/plot03_Silhouette.pdf")
pdf(file = f, width = 8, height = 6)
par(mfrow = c(2,2) )
####
for(pop in pops){
  sil = kmeans[[pop]]$Silhouette
  bestk = kmeans[[pop]]$BestK
  ###
  plot(c(0,sil), type = "b", pch = 19, frame = FALSE,
     main = paste0(pop),
     ylab  = "Avg Silhouette Width",
      xlab = "k")
  abline(v = bestk , lty =2)
}
dev.off()
```


## plot PCs for best K

```{r, fig.width = 10, fig.height = 9}
f = paste0(results_dir, "figures/plot04_continental_kmeans_PCs.pdf")
pdf(f, width = 8, height = 9)

par(mfrow = c(4,3), oma = c(2,1,2,0), mar = c(5,5,2,1))
for(i in 1:length(pops)){
  ## grab the PCs
  pop = names(mydata)[i]
  varexp = signif( mydata[[i]]$varexp[1:20] , d = 3)
  plotdata = mydata[[i]]$pcs
  ## how many clusters ?
  bestk = kmeans[[pop]]$BestK
  ## grab the clusters
  kcluster = kmeans[[pop]]$K[[bestk]]$cluster
  ## Set up the pc data
  plotdata = plotdata[, c("PC1","PC2","PC3","PC4","pop")]
  keepukbb = which(plotdata$pop == "UKBB")
  plotdata = plotdata[keepukbb, ]
  ## Add the K pops
  plotdata$k = as.factor( paste0( "K", kcluster ) )
  ## Add Color
  cols = RColorBrewer::brewer.pal( length(levels(plotdata$k)) , "Set1")
  plotdata$pcol = cols[plotdata$k]
  ## Add Shapes
  shapes = c(15,17,18,19,1,2,3,5,4)[ 1:length(levels(plotdata$k)) ]
  plotdata$ppch = shapes[plotdata$k]
  
  #### The plots
  for(j in 2:4){
    plot(plotdata$PC1, plotdata[,j],
         col = plotdata$pcol,
         pch = plotdata$ppch,
         cex = plotdata$pcex,
         xlab = paste0("PC1 varexp = ", varexp[1], "%" ),
         ylab = paste0("PC",j," varexp =  ", varexp[j], "%" ),
         main = pop)
    abline(h = 0, lty = 2, col = "grey", lwd = 0.75)
    abline(v = 0, lty = 2, col = "grey", lwd = 0.75)
    if(j == 3){
      ## identify the proper color for each pop
      p = levels(plotdata$k)
      cols = sapply(p, function(x){ 
        w = which(plotdata$k == x)[1]; return(plotdata$pcol[w]) 
        })
      ### plot the legend
      if(pop == "AFR"){
        legend("topleft", legend = levels(plotdata$k), 
             pch = 20, col = cols, pt.cex = 2.5, cex = 0.75, bg = NULL)
      } else {
        legend("topright", legend = levels(plotdata$k), 
             pch = 20, col = cols, pt.cex = 2.5, cex = 0.75, bg = NULL)  
      }
      
      }
  }
}

dev.off()


```
## Plot EUR with a K of 6

```{r, fig.width = 10, fig.height = 3}
par(mfrow = c(1,3), oma = c(2,1,2,0), mar = c(5,5,2,1))
  i = 2
  ## grab the PCs
  pop = "EUR"
  varexp = signif( mydata[[i]]$varexp[1:20] , d = 3)
  plotdata = mydata[[i]]$pcs
  ## how many clusters ?
  bestk = kmeans[[pop]]$BestK
  bestk = 6
  ## grab the clusters
  kcluster = kmeans[[pop]]$K[[bestk]]$cluster
  ## Set up the pc data
  plotdata = plotdata[, c("PC1","PC2","PC3","PC4","pop")]
  keepukbb = which(plotdata$pop == "UKBB")
  plotdata = plotdata[keepukbb, ]
  ## Add the K pops
  plotdata$k = as.factor( paste0( "K", kcluster ) )
  ## Add Color
  cols = RColorBrewer::brewer.pal( length(levels(plotdata$k)) , "Set1")
  plotdata$pcol = cols[plotdata$k]
  ## Add Shapes
  shapes = c(15,17,18,19,1,2,3,5,4)[ 1:length(levels(plotdata$k)) ]
  plotdata$ppch = shapes[plotdata$k]
  
  #### The plots
  for(j in 2:4){
    plot(plotdata$PC1, plotdata[,j],
         col = plotdata$pcol,
         pch = plotdata$ppch,
         cex = plotdata$pcex,
         xlab = paste0("PC1 varexp = ", varexp[1], "%" ),
         ylab = paste0("PC",j," varexp =  ", varexp[j], "%" ),
         main = pop)
    abline(h = 0, lty = 2, col = "grey", lwd = 0.75)
    abline(v = 0, lty = 2, col = "grey", lwd = 0.75)
    if(j == 3){
      ## identify the proper color for each pop
      p = levels(plotdata$k)
      cols = sapply(p, function(x){ 
        w = which(plotdata$k == x)[1]; return(plotdata$pcol[w]) 
        })
      ### plot the legend
      if(pop == "AFR"){
        legend("topleft", legend = levels(plotdata$k), 
             pch = 20, col = cols, pt.cex = 2.5, cex = 0.75, bg = NULL)
      } else {
        legend("topright", legend = levels(plotdata$k), 
             pch = 20, col = cols, pt.cex = 2.5, cex = 0.75, bg = NULL)  
      }
      
      }
  }


```



```{r, fig.width = 8, fig.height= 2}

KmeansPCA_Plots = lapply(pops, function(pop){
  ####################################
  ## Define the working plot data set
  ####################################
  plotdata = mydata[[pop]]$pcs
  ## how many clusters ?
  bestk = kmeans[[pop]]$BestK
  if(pop == "EUR"){
    bestk = 6
  }
  ## grab the clusters
  kcluster = kmeans[[pop]]$K[[bestk]]$cluster
  ## Set up the pc data
  plotdata = plotdata[, c("PC1","PC2","PC3","PC4","pop")]
  keepukbb = which(plotdata$pop == "UKBB")
  plotdata = plotdata[keepukbb, ]
  ## Add the K pops
  plotdata$k = as.factor( paste0( "K", kcluster ) )
  
  l = length( levels(plotdata$k) )
  plot_colors = c( RColorBrewer::brewer.pal(9, "Set1") )
  
  ####################################
  ## Define variance Explained
  ####################################
  varexp = mydata[[pop]]$varexp[1:10]
  varexp = round(varexp, d = 3)
  
  ################################
  ## PLOT
  ################################
  myplots = lapply(2:4, function(i){
    pc = paste0("PC", i)
    plotdata %>% ggplot( aes_string(x = "PC1", y = pc)  ) +
    geom_point( aes(color = k), shape = 19 ) + 
    scale_color_manual( values = plot_colors, drop = FALSE ) +
    theme_bw() +
    # scale_fill_manual( values = UNregions$colors, drop = FALSE ) +
    labs(color = "Population", 
         x = paste0("PC1  ", varexp[1], "%"),
         y = paste0("PC", i, "  ", varexp[i], "%")) +
    #guides(color=guide_legend(ncol=2)) +
    guides(color=guide_legend(ncol=1,  override.aes = list(size=4) ) )
  })
  
  p = ggpubr::ggarrange(plotlist = myplots,
                    ncol=3, nrow=1,
                    common.legend = TRUE,
                    legend="right")
  myplots$combined = p
  return(myplots)
})
names(KmeansPCA_Plots) = pops


####################
## Save the ggplot
## objects to 
## file
####################
f = paste0(results_dir, "figures/Kmeans_PCA_Plots_EUR_K6.Rdata")  
save(KmeansPCA_Plots, file = f)


```

```{r, fig.width = 10, fig.height = 4}
myplots = list(KmeansPCA_Plots[["AFR"]]$combined,
               KmeansPCA_Plots[["EUR"]]$combined,
               KmeansPCA_Plots[["SAS"]]$combined,
               KmeansPCA_Plots[["EAS"]]$combined)

f = paste0(results_dir, "figures/plot04_continental_kmeans_PCs_ggplot_v2.pdf")  
pdf(f, width = 11, height = 12)
ggpubr::ggarrange(plotlist = myplots,
                    ncol=1, nrow=4,
                    labels = c("A","B","C", "D"))
dev.off()
```



# Set Up smartpca Fst runs

-- copy the sample id files for smartrel and smartpca over to project portal

## Read in and edit the smartpca ID files

```{r}
dd = paste0(results_dir, "smartrel/")
files = list.files(dd)

ids = lapply(pops, function(pop){
  f = paste0(dd, "ukb_1kg_",pop,".IDS")
  d = read.table(f, header = FALSE, as.is = TRUE, sep = " ")
})
names(ids) = pops
```

## Add the Kpop ids

```{r}
for(pop in pops){
  ##################
  ## sample ID file
  ##################
  d = ids[[pop]] 
  ##################
  ## PCdata file
  ##################
  pcdata = mydata[[pop]]$pcs
  ## keep just the UKBB samples
  w = grep("UKBB", pcdata$pop)
  pcdata = pcdata[w, ]
  ##################
  ## Kmeans Data
  ##################
  k = kmeans[[pop]]
  bestk = as.numeric( k$BestK )
  if(pop == "EUR"){bestk = 6}
  kids = k$K[[bestk]]$cluster
  kids = paste0("K", kids)
  ## add Kids to pcdata
  names(kids) = pcdata$ID
  ## replace UKBB IDs with K-ids
  m = match(d[,1], names(kids))
  d$Kid = kids[m]
  w = which( is.na( d$Kid ) )
  d$Kid[w] = d[w,3]
  d = d[, -3]
  ##
  popnames =  unique(d[,3]) 
  w = grep("UKBB", popnames)
  if(length(w)>0){popnames = popnames[-w]}
  popnames = data.frame(names = sort(popnames) )
  ### WRITE 2 FILE
  f = paste0(results_dir, "kmeans/ukb_1kg_",pop,"_Kcluster.IDS")
  write.table(d, file = f, row.names = FALSE, col.names = FALSE, sep = " ", quote = FALSE)
  ##
  f = paste0(results_dir, "smartpca_Fst/", pop,"_popIDs2run.txt")
  write.table(popnames, file = f, row.names = FALSE, col.names = FALSE, sep = " ", quote = FALSE)
  }
 
```

## draw a random sample of individuals from the EUR sample to derive Fst estimates

```{r}
pop = "AFR"
##################
## sample ID file
##################
d = ids[[pop]] 
##################
## PCdata file
##################
pcdata = mydata[[pop]]$pcs
## keep just the UKBB samples
w = grep("UKBB", pcdata$pop)
pcdata = pcdata[w, ]
##################
## Kmeans Data
##################
k = kmeans[[pop]]
bestk = as.numeric( k$BestK )
if(pop == "EUR"){bestk = 6}
kids = k$K[[bestk]]$cluster
kids = paste0("K", kids)
## add Kids to pcdata
names(kids) = pcdata$ID
## replace UKBB IDs with K-ids
m = match(d[,1], names(kids))
d$Kid = kids[m]
w = which( is.na( d$Kid ) )
d$Kid[w] = d[w,3]
d = d[, -3]
## Randomly Sample the K individuals
k = sort( unique(kids) )
for(i in k){
  w = which(d$Kid == i)
  s = sample(w, 250)
  n = w[!w %in% s]
  d$Kid[n] = paste0(d$Kid[n], "_others")
}
##
f = paste0(results_dir, "kmeans/ukb_1kg_",pop,"_Kcluster_randomsample.IDS")
write.table(d, file = f, row.names = FALSE, col.names = FALSE, sep = " ", quote = FALSE)
  
```



## Fst from smartPCA

```{r}
fst_mats = lapply(pops, function(pop){
  f = paste0(results_dir, "smartpca_Fst/", pop, "_phylip_Fst.txt")
  fmat = read.table(f, header = FALSE, skip = 1)
  rownames(fmat) = fmat[,1]
  fmat = fmat[,-1]
  colnames(fmat) = rownames(fmat)
  return(fmat)
  })
names(fst_mats) = pops
```


```{r, fig.width = 10, fig.height = 6}
library(ape)
dmat = as.dist( as.matrix(  fst_mats[[pop]]  ) )
tree_ape = nj(dmat)
tree = hclust(dmat, method = "ward.D2")
tree2 = hclust(dmat, method = "average")
###
par(mfrow = c(2,1), mar = c(1,5,3,0))
plot(tree, hang = -1, xlab = "")
# plot(tree_ape)
plot(tree2, hang = -1, xlab = "")
```



```{r, fig.width = 6, fig.height = 2}
trees = lapply(pops, function(pop){
  dmat = fst_mats[[pop]] 
  w = which(colnames(dmat) %in% paste0("K", 1:10)  )
  dmat = dmat[w,w]
  dmat = as.dist( as.matrix(  dmat ) )
  # UPGMA cluster
  tree1 = hclust(dmat, method = "average")
  tree2 = hclust(dmat, method = "complete")
  ###
  return(tree2)
})
names(trees) = pops
```


```{r, fig.width = 10, fig.height = 6}
par(mfrow = c(2,2))
for(pop in pops){
  plot( trees[[pop]], hang = -1,
        main = pop, lwd = 2, cex = 2, cex.axis = 2, sub = "", xlab = "", ylab = "Fst distances")
}
```


```{r}
FstDist = sapply(pops, function(pop){
  dmat = fst_mats[[pop]] 
  w = which(colnames(dmat) %in% paste0("K", 1:10)  )
  dmat = dmat[w,w]
  dmat = as.dist( as.matrix(  dmat ) )
  ###
  Fst_mean = mean(dmat)
  Fst_max = max(dmat)
  Fst_min = min(dmat)
  ###
  out = c(Fst_mean, Fst_max, Fst_min)
  names(out) = c("mean","max","min")
  return(out)
})

FstDist
```

```{r}
KG_FstDist = sapply(pops, function(pop){
  dmat = fst_mats[[pop]] 
  w = which(colnames(dmat) %in% paste0("K", 1:10)  )
  dmat = dmat[-w,-w]
  dmat = as.dist( as.matrix(  dmat ) )
  ###
  Fst_mean = mean(dmat)
  Fst_max = max(dmat)
  Fst_min = min(dmat)
  ###
  out = c(Fst_mean, Fst_max, Fst_min)
  names(out) = c("mean","max","min")
  return(out)
})

KG_FstDist
```


```{r}
par(mfrow = c(1,2))
heatmap(KG_FstDist)
heatmap(FstDist)
```

```{r, fig.width = 10, fig.height = 4}
b <- seq(0, 0.02, by = 0.005)
####################################
## UKBB
####################################
x = reshape2::melt(FstDist)
colnames(x) = c("estimate","population","value")
x$estimate = factor(x$estimate, levels = c("max","mean","min") )
x$value = round(x$value, d = 4)
###
P1 = ggplot(x, aes(x = population, y = estimate)) +
  geom_tile(aes(fill = value), color = "white") + 
  geom_text(aes(label=value), color = "black") +
  labs(y = "Fst estimate", x = "continental ancestry group", title = "UKBB k-cluster groups") +
  #scale_fill_viridis_c(option = "B", direction = +1)
  scale_fill_distiller(palette = "YlGnBu", direction = 1, breaks = b, labels = format(b))

####################################
## 1000 Genomes
####################################
x = reshape2::melt(KG_FstDist)
colnames(x) = c("estimate","population","value")
x$estimate = factor(x$estimate, levels = c("max","mean","min") )
x$value = round(x$value, d = 4)
###
P2 = ggplot(x, aes(x = population, y = estimate)) +
  geom_tile(aes(fill = value), color = "white") + 
  geom_text(aes(label=value), color = "black") +
  labs(y = "Fst estimate", x = "continental ancestry group", title = "1000 Genome Populations") +
  #scale_fill_viridis_c(option = "B", direction = +1)
  scale_fill_distiller(palette = "YlGnBu", direction = 1, breaks = b, labels = format(b))

f = paste0(results_dir, "figures/Fst_CAG_estiamtes.pdf")
pdf(f, width = 10, height = 3)
ggpubr::ggarrange(P2, P1, nrow = 1, labels = c("A","B"))
dev.off()

```

